---
title: "Visuals"
author: "Mandy Liesch"
date: "2025-03-10"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



```{r, include=FALSE, echo=FALSE, message=FALSE, warning=FALSE}
library(tidyverse)
library(terra)
library(tidyterra)
library(plotly)
library(sf)
library(ggdendro)
library(dendextend)

options(scipen=999)
options(digits=2)

LUFile<-read_csv("s84_land_cover_basin.csv")

GCAM_Basin_ID<-read_csv("GCAM_Basin_IDs.csv")

GCAM_Boundaries<-st_read("reg_basin_boundaries_moirai_landcells_3p1_0p5arcmin.shp")

GCAM_Boundaries[is.na(GCAM_Boundaries)] <- "Danube"

Basin_Codes<-read_csv("basinCodes.csv")

categories<-read_csv("categories.csv")

```

```{r, echo=FALSE, message=FALSE, warning=FALSE, echo=FALSE}
total_Land_Use<-LUFile  %>% 
  separate(Variable, c("Land", "LandUse", "Status")) %>%
  select(-c("Model", "Unit"))  %>%
  filter(Land == "LandCover") %>%
  filter(LandUse== "Total") %>%
  filter(scenario=="REF") %>%
  select(c("region", "riverBasin", "2010")) %>%
  rename(Area = "2010")


new_Shock_LU<- LUFile  %>% 
  separate(Variable, c("Land", "LandUse", "Status")) %>%
  select(-c("Model", "Unit"))  %>%
  filter(Land == "LandCover") %>%
  filter(LandUse!= "Total") %>%
  filter(Status != "Total"|is.na(Status)) %>%
  mutate(Status = replace_na(Status, "")) %>%
  unite(LandCover, Status, LandUse, sep="")

var1 <- c("2060_P")
var2 <- c("2010_P")

merged_data<-merge(new_Shock_LU, total_Land_Use, by=c("region", "riverBasin")) %>%
     mutate_at(vars("2010":"2060") , funs(P = ./Area * 100)) %>%
     mutate(PerDif= cur_data()[[var1]] - cur_data()[[var2]])

geo_key_merge<-merge(GCAM_Basin_ID, merged_data, by.x="GLU_Name", by.y="riverBasin", all=TRUE)

```

Total Changes 

```{r, message=FALSE, warning=FALSE, echo=FALSE}
Tot_Change <- geo_key_merge %>%
     filter(PerDif > -1) %>%
      group_by(GCAM_Basin_Name, Basin_ID, region, scenario) %>%
     summarize(sum_positive = sum(PerDif))

Tot_Change_Ref<-Tot_Change %>%
  filter(scenario == "REF") 

GCAM_Map_Data<-merge(GCAM_Boundaries, Tot_Change_Ref, by.x=c("basin_nm", "reg_nm"), by.y=c("GCAM_Basin_Name", "region"), all.x=TRUE)

totChange<-ggplot(GCAM_Map_Data) +
  geom_spatvector(aes(fill = sum_positive, text = paste("Basin Name:",basin_nm)), color = NA) +
  scale_fill_viridis_c(name="Land Use Changes") 

ggplotly(totChange)

#Remove the corn and soy shock data and all individual years except the percent difference

GCAM_Map_Data_Full<-merge(GCAM_Boundaries, geo_key_merge, by.x=c("basin_nm", "reg_nm"), by.y=c("GCAM_Basin_Name", "region"), all.x=TRUE)

RefScenario<- GCAM_Map_Data_Full %>%
  filter(scenario == "REF") %>%
  as.data.frame %>%
  select(c("basin_nm", "reg_nm", "key", "LandCover", "PerDif")) %>%
  pivot_wider(id_cols = c("basin_nm", "reg_nm", "key"), names_from = LandCover, values_from = PerDif) %>%
  mutate_at(c('UnmanagedForest','UnmanagedPasture', "ManagedForest", "ManagedPasture", "OtherArable", "OtherNotArable", "Grassland", "Cropland", "Shrubland"), ~replace_na(.,0)) %>%
  mutate(totForest= UnmanagedForest + ManagedForest) %>%
   mutate(totArable=OtherArable + Cropland) %>%
  mutate_at(c('key'), factor)


LU_Map_File<-merge(GCAM_Boundaries, RefScenario, by=c("key", "reg_nm", "basin_nm"))

ChangeArable<-ggplot(LU_Map_File) +
  geom_spatvector(aes(fill = totArable, text = paste("Basin Name:",basin_nm)), color = NA) +
  scale_fill_viridis_c(name="Arable Cropland")
  
ggplotly(ChangeArable)

ChangeForest<-ggplot(LU_Map_File) +
  geom_spatvector(aes(fill = totForest, text = paste("Basin Name:",basin_nm)), color = NA) +
  scale_fill_viridis_c(name="Forest")

ggplotly(ChangeForest)


UnmanForest<-ggplot(LU_Map_File) +
  geom_spatvector(aes(fill = UnmanagedForest, text = paste("Basin Name:",basin_nm)), color = NA) +
  scale_fill_viridis_c(name="Unmanaged Forest")

ManForest<-ggplot(LU_Map_File) +
  geom_spatvector(aes(fill = ManagedForest, text = paste("Basin Name:",basin_nm)), color = NA) +
  scale_fill_viridis_c()

Shrubland<-ggplot(LU_Map_File) +
  geom_spatvector(aes(fill = Shrubland, text = paste("Basin Name:",basin_nm)), color = NA) +
  scale_fill_viridis_c(name="Shrubland")

ggplotly(Shrubland)

Cropland<-ggplot(LU_Map_File) +
  geom_spatvector(aes(fill = Cropland, text = paste("Basin Name:",basin_nm)), color = NA) +
  scale_fill_viridis_c()

OtherArable <-ggplot(LU_Map_File) +
  geom_spatvector(aes(fill = OtherArable, text = paste("Basin Name:",basin_nm)), color = NA) +
  scale_fill_viridis_c()

UnmanPast<-ggplot(LU_Map_File) +
  geom_spatvector(aes(fill = UnmanagedPasture, text = paste("Basin Name:",basin_nm)), color = NA) +
  scale_fill_viridis_c("Unmanaged Pasture")

ggplotly(UnmanPast)

Pasture<-ggplot(LU_Map_File) +
  geom_spatvector(aes(fill = ManagedPasture, text = paste("Basin Name:",basin_nm)), color = NA) +
  scale_fill_viridis_c("Managed Pasture")

ggplotly(Pasture)

Grassland<-ggplot(LU_Map_File) +
  geom_spatvector(aes(fill = Grassland, text = paste("Basin Name:",basin_nm)), color = NA) +
  scale_fill_viridis_c(name="Grassland")

ggplotly(Grassland)

ggarrange(Grassland, Shrubland, ChangeForest, UnmanPast)
```

## Prep the Dataset for Analysis 

```{r, message=FALSE, warning=FALSE, echo=FALSE}
#Remove the corn and soy shock data and all individual years except the percent difference

GCAM_Map_Data_Full<-merge(GCAM_Boundaries, geo_key_merge, by.x=c("basin_nm", "reg_nm"), by.y=c("GCAM_Basin_Name", "region"), all.x=TRUE)

RefScenario<- GCAM_Map_Data_Full %>%
  filter(scenario == "REF") %>%
  as.data.frame %>%
  select(c("basin_nm", "reg_nm", "key", "LandCover", "PerDif")) %>%
  pivot_wider(id_cols = c("basin_nm", "reg_nm", "key"), names_from = LandCover, values_from = PerDif) %>%
  mutate_at(c('UnmanagedForest','UnmanagedPasture', "ManagedForest", "ManagedPasture", "OtherArable", "OtherNotArable", "Grassland", "Cropland", "Shrubland"), ~replace_na(.,0)) %>%
   mutate(Forest= UnmanagedForest + ManagedForest) %>%
  mutate(Crops=OtherArable + Cropland) %>%
  select(-c('UnmanagedForest', "ManagedForest", "OtherArable", "Cropland")) %>%
  mutate_at(c('key'), factor)


LU_Map_File<-merge(GCAM_Boundaries, RefScenario, by=c("key", "reg_nm", "basin_nm"))

totalNA<-RefScenario %>% summarise_all(~ sum(is.na(.)))

cormat <- cor(RefScenario %>% keep(is.numeric))



p.mat <- cor_pmat(RefScenario %>% keep(is.numeric))


col <- colorRampPalette(c("#BB4444", "#EE9988", "#FFFFFF", "#77AADD", "#4477AA"))

corrplot(cormat, method="color", col=col(200),  
         type="upper", order="hclust", 
         addCoef.col = "black", # Add coefficient of correlation
         tl.col="black", tl.srt=45, #Text label color and rotation
         # Combine with significance
         p.mat = p.mat,  
         # hide correlation coefficient on the principal diagonal
         diag=FALSE 
         )

```


```{r, message=FALSE, warning=FALSE, echo=FALSE}
Basin_Codes$Notes<-NULL


Cat_Merge<-merge(LU_Map_File, Basin_Codes, by="key", all=TRUE)

Cat_Merge$VISGROUP<-as.factor(Cat_Merge$VISGROUP)

LossesMap <-ggplot(Cat_Merge %>% filter(VISGROUP != 9)) +
  geom_spatvector(aes(fill = VISGROUP, text = paste("Basin Name:",basin_nm)), color = NA) +
   geom_sf(data = Cat_Merge, aes(fill = NA, text = paste("Basin Name:",basin_nm)),color = "grey50")+ 
  scale_fill_viridis_d()

ggplotly(LossesMap)

```

```{r}

New_Cat_Merge<-merge(LU_Map_File, categories, by="key", all=TRUE) 

New_Cat_Merge$LUCategory<-as.factor(New_Cat_Merge$LUCategory)

New_Cat_Merge<-New_Cat_Merge %>%
    mutate(Group = recode(LUCategory, '1' = 'No Change', '2' = 'One Gain/One Loss', '3' =  'One Gain/Two Loss', '4' = 'One Gain/Multiple Loss', '5'='Two Gains/One Loss', '6'='Two Gains/Two Plus Losses' ))


New_Cat_Merge$labels <- paste(
  "Basin:", New_Cat_Merge$basin_nm,"<br>",
                              "Region:", New_Cat_Merge$reg_nm,"<br>",
                              "Total Changes:", sprintf("%0.2f", as.numeric(New_Cat_Merge$...8)),"<br>",
                              "Total Cropland:", sprintf("%0.2f", as.numeric(New_Cat_Merge$totArable)),"<br>",
                              "Unmanaged Pasture:",sprintf("%0.2f", as.numeric(New_Cat_Merge$UnmanagedPasture)),"<br>",
                              "Grassland:",sprintf("%0.2f", as.numeric(New_Cat_Merge$Grassland.x)),"<br>",
                              "Total Forest:",sprintf("%0.2f", as.numeric(New_Cat_Merge$totForest)),"<br>",                   
                              "Shrubland:",sprintf("%0.2f", as.numeric(New_Cat_Merge$Shrubland.x))
)



CompositeBasinFigure<-ggplot(New_Cat_Merge) +
      geom_spatvector(data=New_Cat_Merge %>% filter(Category == "C"), aes(fill = Group, text = labels)) +
    geom_spatvector(data=New_Cat_Merge %>% filter(Category == "B"), aes(fill = Group, text = labels), alpha=0.5)+
      geom_spatvector(data=New_Cat_Merge %>% filter(Category == "A"), aes(fill = Group, text = labels), alpha=0.2)


l<-ggplotly(CompositeBasinFigure) %>%
  layout(list(hoverformat = '.2f'))

htmlwidgets::saveWidget(l, "basintypes.html")


ggplotly(l)
```

```{r}
CategoryFigure1to1<-ggplot(New_Cat_Merge) +
      geom_spatvector(data=New_Cat_Merge %>% filter(LUCategory == 2), aes(fill = BroadClass)) +
      geom_spatvector(data=New_Cat_Merge, fill=NA)

CategoryFigure2to1<-ggplot(New_Cat_Merge) +
    geom_spatvector(data=New_Cat_Merge %>% filter(LUCategory == 5), aes(fill = BroadClass)) +
      geom_spatvector(data=New_Cat_Merge, fill=NA)
                    
CategoryFigure1to2<-ggplot(New_Cat_Merge) +
    geom_spatvector(data=New_Cat_Merge %>% filter(LUCategory == 3), aes(fill = BroadClass)) +
      geom_spatvector(data=New_Cat_Merge, fill=NA)
                    
CategoryFigure2to2<-ggplot(New_Cat_Merge) +
    geom_spatvector(data=New_Cat_Merge %>% filter(LUCategory == 6), aes(fill = BroadClass)) +
      geom_spatvector(data=New_Cat_Merge, fill=NA)               

CategoryFigure3to1<-ggplot(New_Cat_Merge) +
    geom_spatvector(data=New_Cat_Merge %>% filter(LUCategory == 4), aes(fill = BroadClass)) +
      geom_spatvector(data=New_Cat_Merge, fill=NA)  
