---
title: "Visuals"
author: "Mandy Liesch"
date: "2025-03-10"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



```{r}
library(tidyverse)
library(terra)
library(plotly)

LUFile<-read_csv("s84_land_cover_basin.csv")

GCAM_Basin_ID<-read_csv("GCAM_Basin_IDs.csv")

GCAM_Boundaries<-terra::vect("reg_basin_boundaries_moirai_landcells_3p1_0p5arcmin.shp")
```

```{r}
total_Land_Use<-LUFile  %>% 
  separate(Variable, c("Land", "LandUse", "Status")) %>%
  select(-c("Model", "Unit"))  %>%
  filter(Land == "LandCover") %>%
  filter(LandUse== "Total") %>%
  filter(scenario=="REF") %>%
  select(c("region", "riverBasin", "2010")) %>%
  rename(Area = "2010")


new_Shock_LU<- LUFile  %>% 
  separate(Variable, c("Land", "LandUse", "Status")) %>%
  select(-c("Model", "Unit"))  %>%
  filter(Land == "LandCover") %>%
  filter(LandUse!= "Total") %>%
  filter(Status != "Total"|is.na(Status)) %>%
  mutate(Status = replace_na(Status, "")) %>%
  unite(LandCover, Status, LandUse, sep="")

var1 <- c("2060_P")
var2 <- c("2010_P")

merged_data<-merge(new_Shock_LU, total_Land_Use, by=c("region", "riverBasin")) %>%
     mutate_at(vars("2010":"2060") , funs(P = ./Area * 100)) %>%
     mutate(PerDif= cur_data()[[var1]] - cur_data()[[var2]])

geo_key_merge<-merge(GCAM_Basin_ID, merged_data, by.x="GLU_Name", by.y="riverBasin", all=TRUE)

ManPast_Ref_Merge<- geo_key_merge %>%
  filter(scenario == "REF") %>%
  filter(LandCover == "ManagedPasture")

GCAM_Map_Data<-merge(GCAM_Boundaries, ManPast_Ref_Merge, by.x=c("basin_nm", "reg_nm"), by.y=c("GCAM_Basin_Name", "region"), all.x=TRUE)

ggplot(GCAM_Map_Data) +
  geom_spatvector(aes(fill = PerDif), color = NA) +
  scale_fill_viridis_d()



