---
title: "Global Land Use"
author: "Mandy Liesch"
date: "2025-03-01"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)

LUFile<-read_csv("s84_land_cover_basin.csv")

```


```{r}

new_Shock_LU<- LUFile  %>% 
  separate(Variable, c("Land", "LandUse", "Status")) %>%
  dplyr::select(-c("Model", "Unit"))  %>%
  filter(Land == "LandCover") %>%
  filter(LandUse!= "Total") %>%
  filter(Status != "Total"|is.na(Status)) %>%
  mutate(Status = replace_na(Status, "")) %>%
  unite(LandCover, Status, LandUse, sep="")


wide_sum_cat_LU<- new_Shock_LU %>% 
  dplyr::select(-c("Land"))  %>%
  pivot_longer(!c(scenario, region, riverBasin, LandCover), names_to=c("year"), values_to = "area") %>%
  mutate(year = as.numeric(year)) %>% 
  pivot_wider(names_from = scenario, values_from = c(area)) %>%
  mutate(DifCornShock = CORN1BG - REF) %>%
  mutate(DifSoyShock = SOY1BG - REF)

futRefLand.plt<-ggplot(wide_sum_cat_LU %>% 
                              filter(region=="WLD")) +
  geom_line(mapping=aes(x=year, y=REF, color=LandCover), lwd=2) +
  scale_color_manual(values=c("#FED439FF", "#46732EFF", "green", "#FD8CC1FF", "#D2AF81FF", "#8A9197FF", "#F05C3BFF","#075149FF", "#370335FF"),  name="Land Use", labels=c("Cropland", "Grassland", "Managed Forest", "Managed Pasture", "Other Arable Land","Other Not Arable", "Shrubland", "Unmanaged Forest", "Unmanaged Pasture")) +
  #ggtitle("Future Net Emissions") +
  #Add the theme elements
  #theme(plot.title = element_text(hjust = 0.5)) +
  #Add the labels
  labs( x='Year', y='Area (Mha)', linetype="Shock Type") 

futShocksLand.plt<-ggplot(wide_sum_cat_LU %>% 
                              filter(region=="WLD")) +
  geom_line(mapping=aes(x=year, y=DifCornShock, color=LandCover, linetype="Corn"), lwd=2) +
  geom_line(mapping=aes(x=year, y=DifSoyShock, color=LandCover, linetype="Soy"),  lwd=2) +
  scale_color_manual(values=c("#FED439FF", "#46732EFF", "green", "#FD8CC1FF", "#D2AF81FF", "#8A9197FF", "#F05C3BFF","#075149FF", "#370335FF"),  name="Land Use", labels=c("Cropland", "Grassland", "Managed Forest", "Managed Pasture", "Other Arable Land","Other Not Arable", "Shrubland", "Unmanaged Forest", "Unmanaged Pasture")) +
  #ggtitle("Future Net Emissions") +
  #Add the theme elements
  #theme(plot.title = element_text(hjust = 0.5)) +
  #Add the labels
  labs( x='Year', y='Area (Mha)', linetype="Shock Type") 



futRefLand.plt

  