---
title: "Global Net Carbon Emissions"
author: "Mandy Liesch"
date: "2025-02-28"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

setwd("C:/Users/ALIESCH/OneDrive - Environmental Protection Agency (EPA)/Desktop/gcamsoilc")
```

```{r}
LUColors <- c("UrbanLand"="#C80813FF", "Tundra"="#71D0F5FF", 
              "Rock/Ice/Desert"="#8A9197FF", "Cropland" = "#FED439FF",
              "HarvCropLand"="#FED439FF", "OtherArableLand" = "#D2AF81FF", 
              "Pasture"= "#FD8CC1FF", "UnmanagedPasture"="#370335FF", 
              "Grassland" = "#46732EFF", "Forest" = "#075149FF", 
              "Shrubland" = "#F05C3BFF", "otherNotArable"="#8A9197FF")

# Vegetative Based Color Pallet
vegCEmissions<-c("#46732EFF")

ModelColors<- c("FAO"="#370335FF", "Houghton" = "#FED439FF", 
                "GCAM" = "#FED439FF", "HWSD"="#1A9993FF", "SG17" = "#197EC0FF",
                "SG20"="#F05C3BFF")



```


```{r}
library(tidyverse)

#Create a series loop to read in all of the data

#Create a list of all .csv files in the output folder
files <- list.files(path="World", pattern = "\\.csv$", 
                    full.names = TRUE)

#Initialize an empty list to store data frames
data_list <- list()

# Loop over each file and read it into a data frame
for (file in files) {
  data <- read.csv(file)
  data_list[[file]] <- data
}

# Combine all data frames into one
combined_shock_global_data <- bind_rows(data_list)

new_Shock_Emiss<- combined_shock_global_data  %>% 
  separate(Variable, c("Emissions", "CO2", "LUC", "Source")) %>%
  dplyr::select(-c("series", "Emissions", "CO2", "LUC", "run", "Unit", "model_run_date", "script_run_date", "comments"))  %>%
  dplyr::select(Model:'X2060') %>%
  rename("LandUse" = "landType") %>%
  filter(Source == "Total")

#Reclassify land use types into broader categories
cat_LU<- new_Shock_Emiss %>% 
  mutate(Category = case_when(
  #Cropland category  
  LandUse %in% 
    #Biomass Crops
    c("biomassGrass", "biomassTree",  
    #Fodder Crops
      "FodderGrass", "FodderHerb", 
    #Grain Crops
      "Corn",  "Soybean", "Wheat", "OtherGrain", "Rice",
    #Oil Crops
      "OilCrop", "Rapeseed", "PalmFruit",
    #Other Crops
      "FiberCrop", "MiscCrop", "RootTuber",  
    #Sugar Crops
    "SugarCrop")
    #Land Use Name
      ~ "HarvCropLand",
    #Other Arable Land
    LandUse == "OtherArableLand" ~"OtherArableLand",
    #Forest categories
    LandUse %in% c("Forest", "UnmanagedForest", "ProtectedUnmanagedForest")
                ~ "Forest",
    #Grassland categories
    LandUse %in% c("Grassland", "ProtectedGrassland") ~ "Grassland", 
    #Pasture category
    LandUse=="Pasture" ~ "Pasture",
    #Unamanged pasture category
    LandUse %in% c("ProtectedUnmanagedPasture", "UnmanagedPasture") 
              ~ "UnmanagedPasture",
    #Shrubland category
    LandUse %in% c("Shrubland", "ProtectedShrubland") ~ "Shrubland",
    #Other categories
    LandUse=="UnmanagedLand" ~ "UnmanagedLand",
    LandUse=="RockIceDesert" ~ "RockIceDesert",
    LandUse=="Tundra" ~ "Tundra",
    LandUse=="UrbanLand" ~ "UrbanLand",
    #Default Return Statement
    TRUE~NA_character_
  )
)

#Reclassify land use types into broader categories
reg_LU<- cat_LU %>% 
  mutate(Reg8 = case_when(
  #Cropland category  
  region %in% 
    #Africa
    c("Africa_Eastern", "Africa_Southern", "Africa_Western", "Africa_Northern", "South Africa") 
      ~ "AFR",
    #Other Arable Land
   region == "Brazil" ~"BRA",
    #Pasture category
    region=="China" ~ "CHN",
    #Unamanged pasture category
    region %in% c("EU-12", "EU-15") 
              ~ "EUR",
    #Shrubland category
    region %in% c("Canada", "Europe_Eastern", "Europe_Non_EU", "European Free Trade Association", "Middle East", "Russia") ~ "ROW",
    #Other categories
    region=="USA" ~ "USA",
    region=="WLD" ~ "World",
    region %in% c("Australia_NZ", "Central Asia", "India", "Indonesia", "Japan", "Pakistan", "South Asia", "South Korea", "South Asia", "Southeast Asia",  "Taiwan" ) ~ "XAS", 
   region %in% c("Argentina", "Central America and Caribbean", "Colombia", "Mexico", "South America_Northern", "South America_Southern") ~ "XLM",
    #Default Return Statement
    TRUE~NA_character_
  )
)


#Reclassify land use types into broader categories
reg_2x_LU<- reg_LU %>% 
  mutate(US_RW = case_when(
  #Cropland category  
  Reg8 %in%
    #Africa
    c("AFR", "BRA", "CHN", "EUR", "ROW", "XAS", "XLM") 
      ~ "NonUSA",
    #Other categories
    Reg8=="USA" ~ "USA",
    Reg8=="World" ~ "World",
    #Default Return Statement
    TRUE~NA_character_
  )
)


#Sum everything up by defined land use category
sum_cat_emiss<- cat_LU %>% 
  group_by(Model, scenario, region, riverBasin, Category) %>% 
  summarise(across(where(is.numeric), sum))



#Sum everything up by defined region
sum_cat_reg_emiss<- reg_LU %>% 
  group_by(Model, scenario, Category, Reg8) %>% 
  summarise(across(where(is.numeric), sum))

#Sum everything up by defined region
sum_cat_NoUSA_emiss<- reg_2x_LU %>% 
  group_by(Model, scenario, Category, US_RW) %>% 
  summarise(across(where(is.numeric), sum))

#Create a long dataframe with land use, years, and area 
wide_sum_cat_emiss<- sum_cat_emiss %>% 
  pivot_longer(!c(Model, scenario, region, riverBasin, Category), names_to=c("year"), values_to = "emissions") %>%
  #Remove the extra X value from the dataset
  mutate(across(c('year'), substr, 2, nchar(year))) %>%
  mutate(year = as.numeric(year)) %>% 
  pivot_wider(names_from = scenario, values_from = c(emissions)) %>%
  mutate(DifCornShock = CORN1BG - REF) %>%
  mutate(DifSoyShock = SOY1BG - REF)


#Create a long dataframe with regions, land use, years, and area 
wide_sum_cat_reg_emiss<- sum_cat_reg_emiss %>% 
  pivot_longer(!c(Model, scenario, Reg8, Category), names_to=c("year"), values_to = "emissions") %>%
  #Remove the extra X value from the dataset
  mutate(across(c('year'), substr, 2, nchar(year))) %>%
  mutate(year = as.numeric(year)) %>% 
  pivot_wider(names_from = scenario, values_from = c(emissions)) %>%
  mutate(DifCornShock = CORN1BG - REF) %>%
  mutate(DifSoyShock = SOY1BG - REF)


#Create a long dataframe with regions, land use, years, and area 
wide_sum_cat_NonUS_emiss<- sum_cat_NoUSA_emiss %>% 
  pivot_longer(!c(Model, scenario, US_RW, Category), names_to=c("year"), values_to = "emissions") %>%
  #Remove the extra X value from the dataset
  mutate(across(c('year'), substr, 2, nchar(year))) %>%
  mutate(year = as.numeric(year)) %>% 
  pivot_wider(names_from = scenario, values_from = c(emissions)) %>%
  mutate(DifCornShock = CORN1BG - REF) %>%
  mutate(DifSoyShock = SOY1BG - REF)
```

```{r}

world_sum_cum_emiss<- wide_sum_cat_emiss %>%
  filter(region == "WLD") %>%
  #filter(Category != "UnmanagedLand") %>%
  filter(year >2015) %>%
  #filter(Model == "FAO"|Model=="Houghton") %>%
  group_by(Model, year) %>% 
  summarise(across(where(is.numeric), sum)) %>%
  group_by(Model) %>% 
  mutate(cumSumCorn=cumsum(DifCornShock)) %>%
  mutate(cumSumSoy=cumsum(DifSoyShock)) %>%
  mutate(cumSumRef=cumsum(REF))
  
#Calculate the net emissions plots by series
futNetEmissions.plt<-ggplot() +
  geom_line(world_sum_cum_emiss, mapping=aes(x=year, y=DifCornShock, color=Model, linetype = "Corn"), lwd=2) +
  geom_line(world_sum_cum_emiss, mapping=aes(x=year, y=DifSoyShock, color=Model, linetype = "Soy"), lwd=2) +
      #Add the custom color manual and labels
      scale_color_manual(values=c(ModelColors, name = "Emissions"), 
                      labels = c("FAO Glosis", "Houghton", "HWSD", "SG17", "SG20")) +
  #Add the title==
  #ggtitle("Future Net Emissions") +
  #Add the theme elements
  #theme(plot.title = element_text(hjust = 0.5)) +
  #Add the labels
  labs( x='Year', y='Total Emissions (Mt)', linetype="Shock Type") 
  


futCumNetEmissions.plt<-ggplot() +
  geom_line(world_sum_cum_emiss, mapping=aes(x=year, y=cumSumCorn, color=Model, linetype = "Corn"), lwd=2) +
  geom_line(world_sum_cum_emiss, mapping=aes(x=year, y=cumSumSoy, color=Model, linetype = "Soy"), lwd=2) +
      #Add the custom color manual and labels
      scale_color_manual(values=c(ModelColors, name = "Emissions"), 
                      labels = c("FAO Glosis", "Houghton", "HWSD", "SG17", "SG20")) +
  #Add the title==
  #ggtitle("Future Net Emissions") +
  #Add the theme elements
  #theme(plot.title = element_text(hjust = 0.5)) +
  #Add the labels
  labs( x='Year', y='Total Emissions (Mt)', linetype="Shock Type") 



world_sum_cum_emiss<- wide_sum_cat_emiss %>%
  filter(region == "WLD") %>%
  filter(Category != "UnmanagedLand") %>%
  filter(year >2015) %>%
  filter(Model == "FAO"|Model=="Houghton"|Model=="HWSD"|Model=="SG17"|Model=="SG20") %>%
  group_by(Model, year) %>% 
  summarise(across(where(is.numeric), sum))
  
futRefEmissions.plt<-ggplot() +
  geom_line(world_sum_cum_emiss, mapping=aes(x=year, y=REF, color=Model), lwd=2) +
      #Add the custom color manual and labels
      scale_color_manual(values=c(ModelColors, name = "Emissions"), 
                      labels = c("FAO Glosis", "Houghton", "HWSD", "SG17", "SG20")) +
  #Add the title==
  #ggtitle("Future Net Emissions") +
  #Add the theme elements
  #theme(plot.title = element_text(hjust = 0.5)) +
  #Add the labels
  labs( x='Year', y='Total Emissions (Mt)', linetype="Shock Type") 

futRefCumEmissions.plt<-ggplot() +
  geom_line(world_sum_cum_emiss, mapping=aes(x=year, y=cumSumRef, color=Model), lwd=2) +
      #Add the custom color manual and labels
      scale_color_manual(values=c(ModelColors, name = "Emissions"), 
                      labels = c("FAO Glosis", "Houghton", "HWSD", "SG17", "SG20")) +
  #Add the title==
  #ggtitle("Future Net Emissions") +
  #Add the theme elements
  #theme(plot.title = element_text(hjust = 0.5)) +
  #Add the labels
  labs( x='Year', y='Total Emissions (Mt)', linetype="Shock Type") 


world_LU_FAO_cum_emiss<- wide_sum_cat_emiss %>%
  filter(region == "WLD") %>%
  filter(Category != "UnmanagedLand") %>%
  filter(year >2015) %>%
  filter(Model == "FAO") 

#Calculate the net emissions plots by series
futFAOLUNetEmissions.plt<-ggplot()+
  geom_line(world_LU_FAO_cum_emiss, mapping=aes(x=year, y=REF, group=Category, color=Category), lwd=2) + 
  #Add the custom color manual and labels
      scale_color_manual(values=c(LUColors, name = "Land Use"), 
                      labels = c("Forest", "Grassland", "Harvested Cropland", "Other Arable Land", "Pasture", "Shrubland", "Tundra", "Unmanaged Pasture", "Urban Land")) +
  ggtitle("FAO Global Land Use Emissions in the Reference Scenario") + 
  labs(y="Emissions (Mt)")
```

```{r}
world_reg_sum_cum_emiss<- wide_sum_cat_reg_emiss %>%
  filter(Reg8 != "World") %>%
  filter(year >2015) %>%
  group_by(Model, Reg8, year) %>% 
  summarise(across(where(is.numeric), sum)) %>%
  group_by(Model, Reg8) %>% 
  mutate(cumSumCorn=cumsum(DifCornShock)) %>%
  mutate(cumSumSoy=cumsum(DifSoyShock)) %>%
  mutate(cumSumRef=cumsum(REF))

futRegNetEmissions.plt<-ggplot() +
  #geom_line(world_reg_sum_cum_emiss, mapping=aes(x=year, y=DifCornShock, color=Model, linetype = "Corn"), lwd=2) +
  geom_line(world_reg_sum_cum_emiss, mapping=aes(x=year, y=REF, color=Model), lwd=2) +
      #Add the custom color manual and labels
      scale_color_manual(values=c(ModelColors, name = "Emissions"), 
                      labels = c("FAO Glosis", "Houghton", "HWSD", "SG17", "SG20")) +
  #Add the title==
  #ggtitle("Future Net Emissions") +
  #Add the theme elements
  #theme(plot.title = element_text(hjust = 0.5)) +
  #Add the labels
  labs( x='Year', y='Total Emissions (Mt)', linetype="Shock Type") 

FW_futRegNet.plt<- futRegNetEmissions.plt + facet_wrap(~Reg8)


futRegCumEmissions.plt<-ggplot() +
  #geom_line(world_reg_sum_cum_emiss, mapping=aes(x=year, y=DifCornShock, color=Model, linetype = "Corn"), lwd=2) +
  geom_line(world_reg_sum_cum_emiss, mapping=aes(x=year, y=cumSumRef, color=Model), lwd=2) +
      #Add the custom color manual and labels
      scale_color_manual(values=c(ModelColors, name = "Emissions"), 
                      labels = c("FAO Glosis", "Houghton", "HWSD", "SG17", "SG20")) +
  #Add the title==
  #ggtitle("Future Net Emissions") +
  #Add the theme elements
  #theme(plot.title = element_text(hjust = 0.5)) +
  #Add the labels
  labs( x='Year', y='Total Emissions (Mt)', linetype="Shock Type") 

FW_futRegCum.plt<- futRegCumEmissions.plt + facet_wrap(~Reg8)


futRegShockCumEmissions.plt<-ggplot() +
  geom_line(world_reg_sum_cum_emiss, mapping=aes(x=year, y=cumSumCorn, color=Model, linetype = "Corn"), lwd=2) +
  geom_line(world_reg_sum_cum_emiss, mapping=aes(x=year, y=cumSumSoy, color=Model, linetype = "Soy"), lwd=2) +
      #Add the custom color manual and labels
      scale_color_manual(values=c(ModelColors, name = "Emissions"), 
                      labels = c("FAO Glosis", "Houghton", "HWSD", "SG17", "SG20")) +
  #Add the title==
  #ggtitle("Future Net Emissions") +
  #Add the theme elements
  #theme(plot.title = element_text(hjust = 0.5)) +
  #Add the labels
  labs( x='Year', y='Total Emissions (Mt)', linetype="Shock Type") 

FW_futRegShockCum.plt<- futRegShockCumEmissions.plt + facet_wrap(~Reg8)
  
```

```{r}
world_NoUSA_sum_cum_emiss<- wide_sum_cat_NonUS_emiss %>%
  filter(US_RW != "World") %>%
  filter(year >2015) %>%
  group_by(Model, US_RW, year) %>% 
  summarise(across(where(is.numeric), sum)) %>%
  group_by(Model, US_RW) %>% 
  mutate(cumSumCorn=cumsum(DifCornShock)) %>%
  mutate(cumSumSoy=cumsum(DifSoyShock)) %>%
  mutate(cumRef=cumsum(REF))

futNonUSANetEmissions.plt<-ggplot() +
  geom_line(world_NoUSA_sum_cum_emiss, mapping=aes(x=year, y=REF, color=Model), lwd=2) +
  #geom_line(world_NoUSA_sum_cum_emiss, mapping=aes(x=year, y=DifCornShock, color=Model, linetype = "Corn"), lwd=2) +
  #geom_line(world_NoUSA_sum_cum_emiss, mapping=aes(x=year, y=DifSoyShock, color=Model, linetype = "Soy"), lwd=2) +
      #Add the custom color manual and labels
      scale_color_manual(values=c(ModelColors, name = "Emissions"), 
                      labels = c("FAO Glosis", "Houghton", "HWSD", "SG17", "SG20")) +
  #Add the title==
  #ggtitle("Future Net Emissions") +
  #Add the theme elements
  #theme(plot.title = element_text(hjust = 0.5)) +
  #Add the labels
  labs( x='Year', y='Total Emissions (Mt)', linetype="Shock Type") 

FW_fut_NonUS_Net.plt<- futNonUSANetEmissions.plt + facet_wrap(~US_RW)

futNonUSACumSumRefEmissions.plt<-ggplot() +
  geom_line(world_NoUSA_sum_cum_emiss, mapping=aes(x=year, y=cumRef, color=Model), lwd=2) +
      #Add the custom color manual and labels
      scale_color_manual(values=c(ModelColors, name = "Emissions"), 
                      labels = c("FAO Glosis", "Houghton", "HWSD", "SG17", "SG20")) +
  #Add the title==
  #ggtitle("Future Net Emissions") +
  #Add the theme elements
  #theme(plot.title = element_text(hjust = 0.5)) +
  #Add the labels
  labs( x='Year', y='Total Emissions (Mt)', linetype="Shock Type") 

FW_fut_NonUS_CumSum_Ref.plt<- futNonUSACumSumRefEmissions.plt + facet_wrap(~US_RW)


futNonUSANetSumEmissions.plt<-ggplot() +
  geom_line(world_NoUSA_sum_cum_emiss, mapping=aes(x=year, y=DifCornShock, color=Model, linetype = "Corn"), lwd=2) +
  geom_line(world_NoUSA_sum_cum_emiss, mapping=aes(x=year, y=DifSoyShock, color=Model, linetype = "Soy"), lwd=2) +
      #Add the custom color manual and labels
      scale_color_manual(values=c(ModelColors, name = "Emissions"), 
                      labels = c("FAO Glosis", "Houghton", "HWSD", "SG17", "SG20")) +
  #Add the title==
  #ggtitle("Future Net Emissions") +
  #Add the theme elements
  #theme(plot.title = element_text(hjust = 0.5)) +
  #Add the labels
  labs( x='Year', y='Total Emissions (Mt)', linetype="Shock Type") 

FW_fut_NonUS_NetSum.plt<- futNonUSANetSumEmissions.plt + facet_wrap(~US_RW)

futNonUSACumSumEmissions.plt<-ggplot() +
  geom_line(world_NoUSA_sum_cum_emiss, mapping=aes(x=year, y=cumSumCorn, color=Model, linetype = "Corn"), lwd=2) +
  geom_line(world_NoUSA_sum_cum_emiss, mapping=aes(x=year, y=cumSumSoy, color=Model, linetype = "Soy"), lwd=2) +
      #Add the custom color manual and labels
      scale_color_manual(values=c(ModelColors, name = "Emissions"), 
                      labels = c("FAO Glosis", "Houghton", "HWSD", "SG17", "SG20")) +
  #Add the title==
  #ggtitle("Future Net Emissions") +
  #Add the theme elements
  #theme(plot.title = element_text(hjust = 0.5)) +
  #Add the labels
  labs( x='Year', y='Total Emissions (Mt)', linetype="Shock Type") 

FW_fut_NonUS_CumSum.plt<- futNonUSACumSumEmissions.plt + facet_wrap(~US_RW)
  